---
title: "Lakes without PWLIDs but have P numbers"
author: "Alene Onion"
date: '`r paste("DRAFT Report Generated on:", format(Sys.Date(),format="%B %d, %Y"))`'
output: html_document
---
#  {.tabset}

to check whether lakes without PWL IDs are still found in Chapter X:
https://www.dec.ny.gov/regs/2485.html

To be clear, this is an analysis of the lakes in our database which are only a third of the total lakes in the NHD layer. 

## Percent

A large percentage of the lakes in our database do not have PWLIDs.
To be clear, this is an analysis of the lakes in our database which are only a third of the total lakes in the NHD layer. 

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

fil<-newdata %>% filter(LAKE_WATERBODY_TYPE=='LAKE') %>% 
  select(LAKE_HISTORY_ID) %>% distinct()
ids<-newdata %>% 
  filter(LAKE_HISTORY_ID %in% c(unique(fil$LAKE_HISTORY_ID))) %>% 
  select(LAKE_HISTORY_ID,LOCATION_PWL_ID,LAKE_POND_NUMBER) %>% 
  distinct() %>% 
  mutate(pwl=ifelse(is.na(LOCATION_PWL_ID),'no','yes'),
         pnum=ifelse(is.na(LAKE_POND_NUMBER),'no','yes')) %>% 
  select(LAKE_HISTORY_ID,LAKE_POND_NUMBER,pwl,pnum) %>% distinct() %>% 
  arrange(LAKE_HISTORY_ID,desc(pwl),desc(pnum)) %>% 
  distinct(LAKE_HISTORY_ID,.keep_all = TRUE)
forplot<-ids %>% 
  group_by(pwl,pnum) %>% 
  summarise(n=n()) %>% 
  ungroup()

library(ggplot2)
print(ggplot(data=forplot, aes(x=pwl, y=n,fill=pnum)) +
  geom_bar(stat="identity")+
    xlab("Does the lake have a PWLID?")+
    ylab("Number of lakes")+
    labs(fill='Does the lake have a P number?'))
```

## PWS

Among Public Water supplies,
without PWLIDs but that do have P numbers,
how many of these P numbers that can be found in Chapter X: https://www.dec.ny.gov/regs/2485.html ?


To be clear, we are assuming that the p number in the the lakes database is accurate and can be searched for in the Chapter X html pages. There is likely human error in both assumptions.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(rvest)
library(tidyverse)

files<- list.files(path="C:/Users/amonion/New York State Office of Information Technology Services/LMAS - PWS/PWS_Prioritization/data.requests/NY-CRR/",pattern=".html")
nfiles<-length(files)

ponds<-read.csv("data.requests/nopwl_pond_ID.csv", na.strings="")
#ponds<-ids %>% filter(pwl=="no",pnum=="yes") %>% distinct()
ponds<-ponds %>% select(LAKE_HISTORY_ID,LAKE_POND_NUMBER) %>% distinct()

ponds<-ponds %>%  filter(!is.na(LAKE_POND_NUMBER)) %>% 
  mutate(LAKE_POND_NUMBER=paste("P ",LAKE_POND_NUMBER,sep="")) %>% distinct() %>% 
  #add first row as test
  add_row(LAKE_HISTORY_ID='test known p number',LAKE_POND_NUMBER="P 88") %>% 
  arrange(desc(LAKE_HISTORY_ID)) %>% 
  mutate(found="no")


for(i in 1:nfiles){
content <- read_html(paste("data.requests/NY-CRR/",files[i],sep=""))
#individual test
grepl(ponds$LAKE_POND_NUMBER[1],content)
ponds<-ponds %>% 
  group_by(LAKE_HISTORY_ID) %>% 
  mutate(found=ifelse(grepl(LAKE_POND_NUMBER,content),paste(found,files[i],sep=""),found)) %>% 
  ungroup()
}

forplot<-ponds %>% 
  mutate(found=ifelse(found=="no","no","yes")) %>% 
  group_by(found) %>% 
  summarise(n=n()) %>% 
  ungroup()

library(ggplot2)
print(ggplot(data=forplot, aes(x=found, y=n)) +
  geom_bar(stat="identity")+
    labs(title="Amonth Public Water Supplies without PWL but with P number",
         x="Is the p number found in Chapter X?",
         y="Number of lakes"))
```

## All Waters

Among all lakes,
without PWLIDs but that do have P numbers,
how many of these P numbers that can be found in Chapter X: https://www.dec.ny.gov/regs/2485.html ?


To be clear, we are assuming that the p number in the the lakes database is accurate and can be searched for in the Chapter X html pages. There is likely human error in both assumptions.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(rvest)
library(tidyverse)

files<- list.files(path="C:/Users/amonion/New York State Office of Information Technology Services/LMAS - PWS/PWS_Prioritization/data.requests/NY-CRR/",pattern=".html")
nfiles<-length(files)

#ponds<-read.csv("data.requests/nopwl_pond_ID.csv", na.strings="")
ponds<-ids %>% filter(pwl=="no",pnum=="yes") %>% distinct()
ponds<-ponds %>% select(LAKE_HISTORY_ID,LAKE_POND_NUMBER) %>% distinct()

ponds<-ponds %>%  filter(!is.na(LAKE_POND_NUMBER)) %>% 
  mutate(LAKE_POND_NUMBER=paste("P ",LAKE_POND_NUMBER,sep="")) %>% distinct() %>% 
  #add first row as test
  add_row(LAKE_HISTORY_ID='test known p number',LAKE_POND_NUMBER="P 88") %>% 
  arrange(desc(LAKE_HISTORY_ID)) %>% 
  mutate(found="no")


for(i in 1:nfiles){
content <- read_html(paste("data.requests/NY-CRR/",files[i],sep=""))
#individual test
grepl(ponds$LAKE_POND_NUMBER[1],content)
ponds<-ponds %>% 
  group_by(LAKE_HISTORY_ID) %>% 
  mutate(found=ifelse(grepl(LAKE_POND_NUMBER,content),files[i],found)) %>% 
  ungroup()
}

forplot<-ponds %>% 
  mutate(found=ifelse(found=="no","no","yes")) %>% 
  group_by(found) %>% 
  summarise(n=n()) %>% 
  ungroup()

library(ggplot2)
print(ggplot(data=forplot, aes(x=found, y=n)) +
  geom_bar(stat="identity")+
    labs(title="Amonth Public Water Supplies without PWL but with P number",
         x="Is the p number found in Chapter X?",
         y="Number of lakes"))
```