---
title: "2023.01.12.PWS.wqs.attainment"
author: "Alene Onion"
date: "2023-01-12"
output: html_document
---

#  {.tabset}

to update assessment status of the PWS

## Assessed PWS

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(tidyverse)
rm(list=setdiff(ls(), c("newdata")))
  # setwd("C:/Users/amonion/New York State Office of Information Technology Services/BWAM - Lakes Database/Current")
  # source("new_database/Reading.LMAS.Data.R")
   setwd("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/TP.Variance/TP.Variance")
   
rm(list=setdiff(ls(), c("newdata")))

```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

temp<-newdata %>% 
  filter(PWS=="YES") %>% 
  select(LAKE_HISTORY_ID,LOCATION_PWL_ID) %>% distinct()
att<-read.csv("2022.Data/2022_lmas_evaluations.csv")
att<-att %>% 
  rename(LOCATION_PWL_ID=seg_id) %>% distinct()
att<-merge(temp,att,by=c('LOCATION_PWL_ID'),all.x=TRUE)

#remove NYC reservoirs
#Remove NYC reservoirs 
att<-att %>% filter(!(LAKE_HISTORY_ID %in%
                        c('1302WES0067','1302BOY0076','1302EAS0089','1302BOG0086',
                          '1302MID0062','1302DIV0083','1302CRO0059',
                          '1302TIT0103','1302CRO0109','1302AMA0050','1302MUS0044A','1302NEW0044',
                          '1702KEN1063','1302GIL0061','1302GLE0074','1302KIR0052','1404CAN0402A',
                          '1403PEP0358A','1306RON0815A','1402NEV0058B','1307ASH0848','1202SCH0638A',
                          '1301JER1042','1702HIL1052','1501STA1011','1301TIO0152','1301SUM0193',
                          '1311THO0274','0905CRA0309','1307KIN0837A','1702BAR1108',
                          '1000CHA0001','0300ONT0000')))

rm(list=setdiff(ls(), c("newdata",'att')))
nopwl<-att %>% filter(is.na(LOCATION_PWL_ID))
att<-att %>% filter(!is.na(LOCATION_PWL_ID))
```

Here are the PWS that have sufficient data already for an assessment (n_years_sampled >=2; n_samples>=6; use== source_of_water_supply)

as a bonus, Lindsey asked for the parameters with exceedances in these waterbodies. For example, Attica Reservoir (0102ATT0020) has exceedances (n_exceedances>0) for these parameters: iron, manganese, ammonia

The output should be:

Lake_ID, SDWIS, parameters with exceedance. 

Because there are multiple SDWIS ids per lake (multiple intakes per lake), we reported this list in two tables:

1. PWS lakes that can be assessed

2. Data from table 1 + SDWIS id 


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

table1_expanded<-att %>% 
  filter(n_years_sampled >=2| n_samples_collected>=6) %>% 
  filter(use== 'source_of_water_supply',
         within_period=='TRUE') %>% distinct()

table1_excursions<-table1_expanded %>% filter(n_exceedances>=1) %>%  
  select(LAKE_HISTORY_ID,LOCATION_PWL_ID, parameter) %>% 
  group_by(LAKE_HISTORY_ID, LOCATION_PWL_ID) %>%
  summarise(excursion_parms = toString(parameter)) %>%
  ungroup() %>%
  mutate(excursions = 'Y') %>% 
  distinct()

table1a_assessment_results<-table1_expanded %>% filter(n_exceedances==0, !LAKE_HISTORY_ID %in% table1_excursions$LAKE_HISTORY_ID) %>%  
  select(LAKE_HISTORY_ID,LOCATION_PWL_ID) %>%
  distinct() %>% 
  mutate(excursion_parms = NA,
         excursions = 'N') %>% 
  bind_rows(table1_excursions)

doh<-read.csv("data.requests/PWS_IDs.csv")
doh<-doh %>%
  select(LOCATION_PWL_ID,LAKE_HISTORY_ID,PWS.ID) %>% distinct() 

table1b_PWS_info<-merge(table1a_assessment_results, doh, by = c('LAKE_HISTORY_ID','LOCATION_PWL_ID'), all.x = T)



# # trying to figure out which lakes overlap with the PWL dataset
# where<-att %>% filter(use== 'source_of_water_supply',
#          within_period=='TRUE', LOCATION_PWL_ID %in% sid$LOCATION_PWL_ID)
# t1<-table1a_assessment_results %>%  filter(LOCATION_PWL_ID %in% where$LOCATION_PWL_ID)
# t2<-table2a_assessment_results %>%  filter(LOCATION_PWL_ID %in% where$LOCATION_PWL_ID)
# t3<-table3_ids %>%  filter(LOCATION_PWL_ID %in% where$LOCATION_PWL_ID)


#rm(list=setdiff(ls(), c("newdata",'att','table1')))

DT::datatable(table1a_assessment_results, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(table1b_PWS_info, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```

## Unassessed PWS with data

Here are the PWS that have data but not enough data for an assessment (lakes with data that aren't in the list above)

The output should be:

Lake_ID, SDWIS, parameters with exceedance. 

Because there are multiple SDWIS ids per lake (multiple intakes per lake), we reported this list in two tables:

1. PWS lakes with data that are unassessed

2. Data from table 1 + SDWIS id 
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#remove impaired
remove<-table1a_assessment_results %>% select(LAKE_HISTORY_ID,LOCATION_PWL_ID) %>% distinct() %>% mutate(remove="yes")
table2_expanded<-merge(remove,att,by=c('LAKE_HISTORY_ID','LOCATION_PWL_ID'),all=TRUE)
table2_expanded<-table2_expanded %>% filter(is.na(remove)) %>% select(-remove) %>% 
  #filter to !is.na records
  filter(!is.na(n_samples_collected),use!='fishing')

table2_excursions<-table2_expanded %>% filter(n_exceedances>=1) %>%  
  select(LAKE_HISTORY_ID,LOCATION_PWL_ID,within_period,use, parameter) %>% 
  group_by(LAKE_HISTORY_ID, LOCATION_PWL_ID,within_period) %>%
  summarise(excursion_parms = toString(parameter)) %>%
  ungroup() %>%
  mutate(within_period=ifelse(within_period=="TRUE","recent","old")) %>%  distinct() %>% 
  spread(within_period,excursion_parms) %>% 
  mutate(excursions = 'Y') %>% 
  distinct()

table2a_assessment_results<-table2_expanded %>% filter(n_exceedances==0, !LAKE_HISTORY_ID %in% table2_excursions$LAKE_HISTORY_ID) %>%  
  select(LAKE_HISTORY_ID,LOCATION_PWL_ID) %>%
  distinct() %>% 
  mutate(old = NA, recent = NA,
         excursions = 'N') %>% 
  bind_rows(table2_excursions)

doh<-read.csv("data.requests/PWS_IDs.csv")

table2b_PWS_info<-merge(table2a_assessment_results, doh, by = c('LAKE_HISTORY_ID','LOCATION_PWL_ID'), all.x= T)

DT::datatable(table2a_assessment_results, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(table2b_PWS_info, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```

## Unassessed PWS without data

Here are the PWS without any data, unassessed

The output should be:

Lake_ID, SDWIS. 

Because there are multiple SDWIS ids per lake (multiple intakes per lake), we reported this list in two tables:

1. Unassessed PWS lakes 

2. Data from table 1 + SDWIS id 

The following lakes were NOT SAMPLED because they were non-community systems: 1501STA1011 1301TIO0152 1301SUM0193 1311THO0274 0905CRA0309

The following lakes were NOT SAMPLED because they were emergency back up systems: 1307KIN0837A

The following lake was NOT SAMPLED because it is inaccessible (located on an island off the coast of Connecticut) 1702BAR1108

also removed lake ontario and lake champlain sites since those are sampled by the GL program
1000CHA0001
0300ONT0000


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}


table3<-att %>% filter(is.na(n_samples_collected)) %>%
  select(LAKE_HISTORY_ID, LOCATION_PWL_ID) %>%
  distinct() 


doh<-read.csv("data.requests/PWS_IDs.csv")

table3sidwis<-merge(table3, doh, by = c('LAKE_HISTORY_ID','LOCATION_PWL_ID'), all.x= T)

DT::datatable(table3, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(table3sidwis, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```

## Unassessed without PWLID

As a reminder, we also have public water supplies that aren't currently on the waterbody inventory list. 
Our current approach is to not sample these for now. We would have to add them to the waterbody inventory to do an assessment.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}


doh<-read.csv("data.requests/PWS_IDs.csv")

nopwl_sdwis<-merge(nopwl, doh, by = c('LAKE_HISTORY_ID','LOCATION_PWL_ID'), all.x= T)

DT::datatable(nopwl, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(nopwl_sdwis, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```