---
title: "PWS_unconfirmed_Prioritization"
author: "Alene Onion"
date: "2/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  {.tabset}

## Data set

unconfirmed public water supplies that are not NYC reservoirs

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#merging lake id with the DOH ID

rm(list=ls())
#read in data
#Script to read in new database
library(tidyverse)
# library(ggmap)
# library(ggrepel)
# library(rgdal)
# library(sp)
# library(spatialEco)
# library(tidyverse)
library(knitr)
#Reading in new database
setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
lake<-read.csv("new_database/L_LAKE.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)

setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")

#removed because more accuate list of PWS is from DOH now in our database
#################################################
#merge with PWS layer in the data selector
#############################
#Now pull SDWIS id
# pws<-readOGR("pws_layer","pws_10_2021") #removed because more accuate list of PWS is from DOH now in our database
# wipwl<-readOGR("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Lakes.Database/data/2020/data.requests/gis.data.selector","dec_wipwl_lakes")
# 
# #change coords to web mercator for the map
# #all of our layers are NAD83 so have to convert to web mercator
# wipwl<-sp::spTransform(wipwl, sp::CRS("+proj=longlat +datum=WGS84 +no_defs"))
# pws<-sp::spTransform(pws, sp::CRS("+proj=longlat +datum=WGS84 +no_defs"))
# 
# #do left join
# pws_merge<-point.in.poly(pws,wipwl)
# 
# #convert attribute table to a data frame
# pws_data<-fortify(pws_merge@data)
# pws_data<-pws_data %>% select(WATER_SY_1,PWL_ID,WBCATGRY) %>% filter(!is.na(PWL_ID)) %>% distinct() %>% rename(PWS.ID=WATER_SY_1,LOCATION_PWL_ID=PWL_ID)
# lmas<-merge(location,pws_data,by=c('LOCATION_PWL_ID'),all.x=TRUE)
# lmas<-lmas %>% select(PWS.ID,LAKE_HISTORY_ID,WBCATGRY,LOCATION_WATERBODY_CLASSIFICATION,LOCATION_PWL_ID) %>% filter(!is.na(PWS.ID)) %>% distinct()
# lake_merge<-lake %>% select(LAKE_HISTORY_ID,LAKE_FIN) %>% distinct()
# lmas<-merge(lmas,lake_merge,by=c('LAKE_HISTORY_ID'),all.x=TRUE)
# 
# #add lake name
# lake<-lake %>% select(LAKE_HISTORY_ID,LAKE_WATERBODY_NAME) %>% distinct()
# lmas<-merge(lmas,lake,by=c('LAKE_HISTORY_ID'),all.x=TRUE)
# 
# #remove NYC reservoirs
# lmas<-lmas %>% 
#   filter(!(LOCATION_PWL_ID %in% c('1702-0059','1302-0025','1306-0003')))
# 
# #compare to those that Rebecca pulled previously
# previous<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/PWS.csv")
# previous<-merge(previous,lmas,by=c('LAKE_HISTORY_ID'),all=TRUE)
# previous<-previous %>% select(-LAKE_WATERBODY_NAME)
# previous<-merge(previous,lake,by=c('LAKE_HISTORY_ID'),all.x=TRUE)
# previous<-previous %>% filter(is.na(PWS.ID))
# 
# rm(list=setdiff(ls(), c("lmas")))

lake<-lake %>% select(LAKE_HISTORY_ID,PWS) %>% distinct() %>% 
  mutate(PWS=toupper(PWS)) %>% 
  filter(PWS=="YES")
locs<-location %>% select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,LOCATION_PWL_ID) %>% distinct()
pws<-merge(lake,locs,by=c('LAKE_HISTORY_ID'),all.x=TRUE)

draft<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/2022_stayCALM_ponded_internal-draft-assessments.csv")
draft<-draft %>% 
  select(seg_id:parameter) %>% 
  distinct() %>% 
  rename(LOCATION_PWL_ID=seg_id) %>% 
  mutate(value=1)
draft<-merge(pws,draft,by=c('LOCATION_PWL_ID'),all.x=TRUE)
#restrict to public water supply use
draft<-draft %>% 
  filter(use=="source_of_water_supply") %>% 
  select(-LOCATION_HISTORY_ID) %>% distinct()



#restrict to those that have confirmed assessments
draft<-draft %>% 
  filter(grepl("unconfirmed",use_assessment)) %>% 
  distinct()

#Remove NYC reservoirs
draft<-draft %>% filter(!(LAKE_HISTORY_ID %in% c('1302WES0067','1302BOY0076','1302EAS0089','1302BOG0086','1302MID0062','1302DIV0083','1302CRO0059',
                                                 '1302TIT0103','1302CRO0109','1302AMA0050','1302MUS0044A','1302NEW0044',
                                                 '1702KEN1063','1302GIL0061','1302GLE0074','1302KIR0052','1404CAN0402A','1403PEP0358A',
                                                 '1306RON0815A','1402NEV0058B','1307ASH0848','1202SCH0638A','1301JER1042',
                                                 '1702HIL1052')))

#add in SDWIS id
# draft<-merge(draft,pws_data,by=c('LOCATION_PWL_ID'),all.x=TRUE) #the merge is incomplete so asking DOH for this data
# locs<-location %>% filter(LOCATION_TYPE=="CENTROID") %>% distinct() %>%
#   select(LOCATION_PWL_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()
# draft<-merge(draft,locs,by=c('LOCATION_PWL_ID'),all.x=TRUE)
# junk<-draft %>% filter(is.na(PWS.ID))
# write.csv(junk,file="junk.rebecca.csv",row.names = FALSE)

#tally results
junk<-draft %>% 
  group_by(use_assessment) %>% 
  summarize(n=n()) %>% 
  ungroup()
library(kableExtra)
kable(junk)
```

Map of unconfirmed PWS with PWL IDs

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
sites<-draft %>% select(LOCATION_PWL_ID,use_assessment)
#Reading in new database
setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")
coordinates<-location %>% filter(LOCATION_TYPE=="CENTROID"|LOCATION_NAME=="CHAUTAUQUA S CSLAP SITE 33.1") %>% select(LOCATION_PWL_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()
sites<-merge(sites,coordinates,by=c('LOCATION_PWL_ID'),all.x = TRUE)

library(ggmap)
library(ggrepel)

nybox<-make_bbox(sites,lon=LOCATION_X_COORDINATE,lat=LOCATION_Y_COORDINATE)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites,aes(x=LOCATION_X_COORDINATE,y=LOCATION_Y_COORDINATE,label=LOCATION_PWL_ID),size=4))



```



## SDWIS Data

merge unconfirmed PWS with PWL IDs with SDWIS ids
(___NOTE___ SDWIS IDs are draft as of 1/27/2022)

Then Rank PWS by population served

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
doh<-read.csv("data.requests/PWS_IDs.csv")
doh<-doh %>%
  select(LOCATION_PWL_ID,LAKE_HISTORY_ID,PWS.ID) %>% distinct() 
draft<-merge(draft,doh,by=c('LAKE_HISTORY_ID','LOCATION_PWL_ID'),all.x=TRUE)


```

Consider those that are:
surface water ('Primary.Source')
PopulationServed.Count

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
lmas<-draft
#read in DWSP2 files
wss<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/water.systems.summary.csv")
lmas<-merge(wss,lmas,by=c('PWS.ID'),all=TRUE)

#too many individual facilities to be useful
# rep<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/facility.report.csv")
# lmas<-merge(rep,lmas,by=c('PWS.ID','PWS.Name','Primary.Source','PWS.Type','Owner.Type'),all=TRUE)

viol<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/Water System Summary viol.csv")
#colnames(viol)[(colnames(viol) %in% colnames(lmas))]
lmas<-merge(lmas,viol,by=c("PWS.ID","PWS.Name","PWS.Type","Primary.Source","Counties.Served","Cities.Served","PopulationServed.Count","Number.of.Facilities"),all.x=TRUE)

lmas<-lmas %>% 
   filter(is.na(Primary.Source)|Primary.Source=="Surface water",
          !is.na(LAKE_HISTORY_ID)) %>% 
  #,          Facility.Activity=="Active"
#          Availability.Description=="Permanent")
  arrange(desc(PopulationServed.Count))


write.csv(lmas,file="PWS.for.TMDL.and.sdwis.csv",row.names=FALSE)
DT::datatable(lmas, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```

## PWS without PWLIDs

Map of PWS that don't have PWL IDs

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
draft2<-pws %>% filter(is.na(LOCATION_PWL_ID)) %>% distinct() %>% mutate(PWS=="yes")
sites<-draft2 %>% select(LAKE_HISTORY_ID,PWS) %>% distinct()
#Reading in new database
setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")
coordinates<-location %>% filter(LOCATION_TYPE=="CENTROID") %>% 
  select(LAKE_HISTORY_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()
sites<-merge(sites,coordinates,by=c('LAKE_HISTORY_ID'),all.x = TRUE)

library(ggmap)
library(ggrepel)

nybox<-make_bbox(sites,lon=LOCATION_X_COORDINATE,lat=LOCATION_Y_COORDINATE)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites,aes(x=LOCATION_X_COORDINATE,y=LOCATION_Y_COORDINATE,label=LAKE_HISTORY_ID),size=4))

doh<-read.csv("data.requests/PWS_IDs.csv")
doh<-doh %>%
  select(LOCATION_PWL_ID,LAKE_HISTORY_ID,PWS.ID) %>% distinct() 
sites<-merge(sites,doh,by=c('LAKE_HISTORY_ID'),all.x=TRUE)
sites<-sites %>% select(LAKE_HISTORY_ID,PWS.ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()


DT::datatable(sites, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
```

