---
title: "PWS_Prioritization"
author: "Alene Onion"
date: "10/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  {.tabset}

## Data set

Here is the joined data set of SDWIS Information joined to waterbody assessments 
NOTE: ASSESSMENT ARE DRAFT AND MUST NOT BE SHARED OUTSIDE THE DEPARTMENT

Unfortunately, this file was too large to share in this format
I will have to email it to you separately

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#merging lake id with the DOH ID

rm(list=ls())
#read in data
#Script to read in new database
library(tidyverse)
library(ggmap)
library(ggrepel)
library(rgdal)
library(sp)
library(spatialEco)
library(tidyverse)
library(knitr)
#Reading in new database
setwd("C:/Users/amonion/New York State Office of Information Technology Services/BWAM - Lakes Database/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
lake<-read.csv("new_database/L_LAKE.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
setwd("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization")

#merge with PWS layer in the data selector
#############################
#Now add counties
pws<-readOGR("pws_layer","pws_10_2021")
wipwl<-readOGR("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Lakes.Database/data/2020/data.requests/gis.data.selector","dec_wipwl_lakes")

#change coords to web mercator for the map
#all of our layers are NAD83 so have to convert to web mercator
wipwl<-sp::spTransform(wipwl, sp::CRS("+proj=longlat +datum=WGS84 +no_defs"))
pws<-sp::spTransform(pws, sp::CRS("+proj=longlat +datum=WGS84 +no_defs"))

#do left join
pws_merge<-point.in.poly(pws,wipwl)

#convert attribute table to a data frame
pws_data<-fortify(pws_merge@data)
pws_data<-pws_data %>% select(WATER_SY_1,PWL_ID,WBCATGRY) %>% filter(!is.na(PWL_ID)) %>% distinct() %>% rename(PWS.ID=WATER_SY_1,LOCATION_PWL_ID=PWL_ID)
lmas<-merge(location,pws_data,by=c('LOCATION_PWL_ID'),all.x=TRUE)
lmas<-lmas %>% select(PWS.ID,LAKE_HISTORY_ID,WBCATGRY,LOCATION_WATERBODY_CLASSIFICATION,LOCATION_PWL_ID) %>% filter(!is.na(PWS.ID)) %>% distinct()

#add lake name
lake<-lake %>% select(LAKE_HISTORY_ID,LAKE_WATERBODY_NAME) %>% distinct()
lmas<-merge(lmas,lake,by=c('LAKE_HISTORY_ID'),all.x=TRUE)

#compare to those that Rebecca pulled previously
previous<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/PWS.csv")
previous<-merge(previous,lmas,by=c('LAKE_HISTORY_ID'),all=TRUE)
previous<-previous %>% select(-LAKE_WATERBODY_NAME)
previous<-merge(previous,lake,by=c('LAKE_HISTORY_ID'),all.x=TRUE)
previous<-previous %>% filter(is.na(PWS.ID))

rm(list=setdiff(ls(), c("lmas")))
```

Filtered to those that are:
1. surface water ('Primary.Source')
2. active ('Facility.Activity')
3. Permanent ('Availability.Description')
Other filters to consider:
Is.Wholesaler
Seller.Treatment.Description ("Filtered","Partially treated by seller","Not treated" )
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#read in DWSP2 files
wss<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/water.systems.summary.csv")
lmas<-merge(wss,lmas,by=c('PWS.ID'),all=TRUE)

rep<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/facility.report.csv")
lmas<-merge(rep,lmas,by=c('PWS.ID','PWS.Name','Primary.Source','PWS.Type','Owner.Type'),all=TRUE)

viol<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/Water System Summary viol.csv")
#colnames(viol)[(colnames(viol) %in% colnames(lmas))]
lmas<-merge(lmas,viol,by=c("PWS.ID","PWS.Name","PWS.Type","Primary.Source","Counties.Served","Cities.Served","PopulationServed.Count","Number.of.Facilities"),all.x=TRUE)

lmas<-lmas %>% filter(!is.na(LAKE_HISTORY_ID)) %>% 
  rename(Previous_Assessment=WBCATGRY) %>% distinct()
#DT::datatable(lmas, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

lmas<-lmas %>% 
  filter(Primary.Source=="Surface water",
         Facility.Activity=="Active",
         Availability.Description=="Permanent")

rm(list=setdiff(ls(), c("lmas")))

write.csv(lmas,file="Merged.assessment.and.sdwis.csv",row.names=FALSE)

```

## Classification and Assessment

PLEASE NOTE the assessment data are DRAFT from Zach and Sarah and SHOULD NOT BE SHARED outside of the agency.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}


draft<-read.csv("2022_stayCALM_ponded_internal-draft-assessments.csv")
draft<-draft %>% 
  select(seg_id:parameter) %>% 
  distinct() %>% 
  rename(LOCATION_PWL_ID=seg_id)
lmas<-merge(lmas,draft,by=c('LOCATION_PWL_ID'),all.x=TRUE)

forplot<-lmas %>%
  select(LOCATION_PWL_ID,class) %>% distinct() %>%
  mutate(value=1,class_simple=substr(class,1,1)) 
print(ggplot(forplot,aes(fill=class,x=class_simple)) +
        geom_bar(psoition="dodge",state="identity") )

```
Now remove class B and C waters because we need to table those for now
And plot the distribution of assessments for the source_water_supply use ONLY

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
lmas<-lmas %>% 
  filter(!(class %in% c('b','b(t)','c'))) %>% distinct()

print(ggplot(lmas %>% filter(use=="source_of_water_supply",use_assessment!='not-applicable') %>%
               select(LOCATION_PWL_ID,use,use_assessment) %>% distinct(),aes(fill=use,x=use_assessment)) +
        geom_bar(psoition="dodge",state="identity") +
        theme(axis.title.x = element_blank(),legend.position = "none",
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))
kable(lmas %>% filter(use=="source_of_water_supply",use_assessment!='not-applicable') %>%
               select(LOCATION_PWL_ID,use,use_assessment) %>% distinct() %>% group_by(use_assessment) %>% summarize(n=n()) %>% ungroup())
```

and plot how many assessments list dissolved oxygen as the parameter of concern. 
Repeat for all major parameters

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#now plot the breakdown of the uses we don't have data for
forplot<-lmas %>% 
  filter(use=="source_of_water_supply") %>% select(LOCATION_PWL_ID,use_assessment,parameter) %>% distinct() %>% 
    mutate(phosphorus=ifelse(grepl('phosphorus',parameter),1,0),
         ph=ifelse((grepl('; ph',parameter)|grepl('ph;',parameter)|parameter=="ph"),1,0),
         dissolved_oxygen=ifelse(grepl('dissolved_oxygen',parameter),1,0),
         chloride=ifelse(grepl('chloride',parameter),1,0),
         ammonia=ifelse(grepl('ammonia',parameter),1,0),
         nitrate_nitrite=ifelse(grepl('nitrate',parameter)|grepl('nitrite',parameter)|grepl('nitrate/nitrite',parameter),1,0),
         metals=ifelse(grepl('iron',parameter)|grepl('manganese',parameter)|grepl('magnesium',parameter)|grepl('copper',parameter),1,0),
         arsenic=ifelse(grepl('arsenic',parameter),1,0),
         sulfate=ifelse(grepl('sulfate',parameter),1,0),
         mercury=ifelse(grepl('mercury',parameter),1,0),
         toxics=ifelse(grepl('dioxin',parameter)|grepl('mirex',parameter)|grepl('pcbs',parameter)|grepl('ddt',parameter)|grepl('chlordane',parameter)|grepl('cadmium',parameter)|grepl('pcbs_other_toxics',parameter)|grepl('phenolic_compounds',parameter),1,0),
         fecal_coliform=ifelse(grepl('fecal_coliform',parameter),1,0),
         solids=ifelse(grepl('silt_sediment',parameter)|grepl('total_dissolved_solids',parameter),1,0),
         problem_species=ifelse(grepl('problem_species',parameter)|grepl('algal_weed_growth',parameter)|grepl('aquatic_weeds',parameter),1,0),
         water_level_flow=ifelse(grepl('water_level_flow',parameter),1,0),
         no_data=ifelse(grepl('No_Data',parameter),1,0)) %>% 
  select(-parameter) %>% 
  gather(parameter_simple,value,-LOCATION_PWL_ID,-use_assessment) %>% 
  filter(value!=0) %>% 
  select(-value) %>% distinct() %>% 
  mutate(use="source_water")

print(ggplot(forplot %>% filter(parameter_simple!='no_data') %>% select(LOCATION_PWL_ID,parameter_simple,use_assessment) %>% distinct(),aes(fill=parameter_simple,x=use_assessment)) +
        geom_bar(position="dodge",state="identity") +
        theme(axis.title.x = element_blank(),
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))

```

## Analysis of SDWIS Data

Should we be including NYC reservoirs? Do we assess these?

Filtered to only those that were unassessed or needed more data
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
lmas2<-lmas %>% 
  filter(use=="source_of_water_supply",
         use_assessment %in% c("IR3_impaired_unconfirmed","IR3_stressed_unconfirmed",
                               "IR3_unassessed","IR3_fully-supported_unconfirmed")) %>% distinct()

```


Populations served (Removing NYC reservoirs)

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#parameters to examine:
#Population.Served.Count
#Service.Connections.Count
#Filtration.Status.Description
#Number.of.Violations

#population served
pops<-lmas2 %>% 
  filter(Population.Served.Count<8271000) %>% 
  select(LOCATION_PWL_ID,Population.Served.Count) %>% distinct() %>% 
  mutate(value='x')
print(ggplot(pops ,aes(y=Population.Served.Count)) +
        geom_boxplot() +
        theme(axis.title.x = element_blank(),legend.position = "none"))
```

Service Connections (again removing NYC reservoirs)

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#parameters to examine:
#Service.Connections.Count
#Filtration.Status.Description
#Number.of.Violations

#population served
pops<-lmas2 %>% 
  filter(Service.Connections.Count<727058) %>% 
  select(LOCATION_PWL_ID,Service.Connections.Count) %>% distinct() %>% 
  mutate(value='x')
print(ggplot(pops ,aes(y=Service.Connections.Count)) +
        geom_boxplot() +
        theme(axis.title.x = element_blank(),legend.position = "none"))
```
Filtration Status
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#parameters to examine:
#Filtration.Status.Description
#Number.of.Violations

#population served
pops<-lmas2 %>% 
  #filter(Service.Connections.Count<727058) %>% 
  select(LOCATION_PWL_ID,Filtration.Status.Description) %>% distinct() %>% 
  mutate(value='x')
print(ggplot(pops,aes(fill=value,x=Filtration.Status.Description)) +
        geom_bar(psoition="dodge",state="identity") +
        theme(axis.title.x = element_blank(),legend.position = "none",
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))

```

Owner Type
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#parameters to examine:
#Filtration.Status.Description
#Number.of.Violations

#population served
pops<-lmas2 %>% 
  #filter(Service.Connections.Count<727058) %>% 
  select(LOCATION_PWL_ID,Owner.Type) %>% distinct() %>% 
  mutate(value='x')
print(ggplot(pops,aes(fill=value,x=Owner.Type)) +
        geom_bar(psoition="dodge",state="identity") +
        theme(axis.title.x = element_blank(),legend.position = "none",
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))

```

Number of Violations
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

#parameters to examine:
#Service.Connections.Count
#Filtration.Status.Description
#Number.of.Violations

#population served
pops<-lmas2 %>% 
  #filter(Service.Connections.Count<727058) %>% 
  select(LOCATION_PWL_ID,Number.of.Violations) %>% distinct() %>% 
  mutate(value='x')
print(ggplot(pops ,aes(y=Number.of.Violations)) +
        geom_boxplot() +
        theme(axis.title.x = element_blank(),legend.position = "none"))
```