---
title: "PWS_Prioritization"
author: "Alene Onion, Charlie Stoll"
date: "2/16/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  {.tabset}

The following tabs list the lakes and streams we intend to screen (Sc.) in 2022/23 as well those that we are considering for intensive sampling (Int).
We are requesting input from DOH on those waterbodies that would be a priority from these lists for intensive sampling.

## Sc. PWS Lakes

Lakes with public water supply intakes that we are intending to screen 2022-23. TO be clear, there are multiple intakes on many of these waterbodies.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#merging lake id with the DOH ID

rm(list=ls())
#read in data
#Script to read in new database
library(tidyverse)
# library(ggmap)
# library(ggrepel)
# library(rgdal)
# library(sp)
# library(spatialEco)
# library(tidyverse)
library(knitr)
#Reading in new database
#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
setwd("C:/Users/amonion/New York State Office of Information Technology Services/BWAM - Lakes Database/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
lake<-read.csv("new_database/L_LAKE.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)

#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")
setwd("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization")


lake<-lake %>% select(LAKE_HISTORY_ID,PWS) %>% distinct() %>% 
  mutate(PWS=toupper(PWS)) %>% 
  filter(PWS=="YES")
locs<-location %>% select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,LOCATION_PWL_ID) %>% distinct()
pws<-merge(lake,locs,by=c('LAKE_HISTORY_ID'),all.x=TRUE)

#draft<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/2022_stayCALM_ponded_internal-draft-assessments.csv")
draft<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/2022_stayCALM_ponded_internal-draft-assessments.csv")
draft<-draft %>% 
  select(seg_id:parameter) %>% 
  distinct() %>% 
  rename(LOCATION_PWL_ID=seg_id) %>% 
  mutate(value=1)
draft<-merge(pws,draft,by=c('LOCATION_PWL_ID'),all.x=TRUE)
#restrict to public water supply use
draft<-draft %>% 
  filter(use=="source_of_water_supply") %>% 
  select(-LOCATION_HISTORY_ID) %>% distinct()



#restrict to those that have confirmed assessments
draft<-draft %>% 
  filter(grepl("unassessed",use_assessment)) %>% 
  distinct()

#Remove NYC reservoirs
draft<-draft %>% filter(!(LAKE_HISTORY_ID %in% c('1302WES0067','1302BOY0076','1302EAS0089','1302BOG0086','1302MID0062','1302DIV0083','1302CRO0059',
                                                 '1302TIT0103','1302CRO0109','1302AMA0050','1302MUS0044A','1302NEW0044',
                                                 '1702KEN1063','1302GIL0061','1302GLE0074','1302KIR0052','1404CAN0402A','1403PEP0358A',
                                                 '1306RON0815A','1402NEV0058B','1307ASH0848','1202SCH0638A','1301JER1042',
                                                 '1702HIL1052')))


#tally results
junk<-draft %>% 
  select(LAKE_HISTORY_ID,use_assessment) %>% 
  mutate(sampling="screening") %>% 
  group_by(sampling) %>% 
  summarize(n=n()) %>% 
  ungroup()
library(kableExtra)
kable(junk)
```



```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
sites<-draft %>% select(LOCATION_PWL_ID,use_assessment)
#Reading in new database
#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
setwd("C:/Users/amonion/New York State Office of Information Technology Services/BWAM - Lakes Database/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")
setwd("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization")
coordinates<-location %>% filter(LOCATION_TYPE=="CENTROID"|LOCATION_NAME=="CHAUTAUQUA S CSLAP SITE 33.1") %>% select(LOCATION_PWL_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()
sites<-merge(sites,coordinates,by=c('LOCATION_PWL_ID'),all.x = TRUE)

library(ggmap)
library(ggrepel)

nybox<-make_bbox(sites,lon=LOCATION_X_COORDINATE,lat=LOCATION_Y_COORDINATE)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites,aes(x=LOCATION_X_COORDINATE,y=LOCATION_Y_COORDINATE,label=LOCATION_PWL_ID),size=4))



```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
doh<-read.csv("data.requests/PWS_IDs.csv")
doh<-doh %>%
  select(LOCATION_PWL_ID,LAKE_HISTORY_ID,PWS.ID) %>% distinct() 
draft<-merge(draft,doh,by=c('LAKE_HISTORY_ID','LOCATION_PWL_ID'),all.x=TRUE)

```

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
lmas<-draft
#read in DWSP2 files
#wss<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/water.systems.summary.csv")
wss<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/water.systems.summary.csv")
lmas<-merge(wss,lmas,by=c('PWS.ID'),all=TRUE)

#too many individual facilities to be useful
# rep<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/facility.report.csv")
# lmas<-merge(rep,lmas,by=c('PWS.ID','PWS.Name','Primary.Source','PWS.Type','Owner.Type'),all=TRUE)

#viol<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/Water System Summary viol.csv")
viol<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/Water System Summary viol.csv")

#colnames(viol)[(colnames(viol) %in% colnames(lmas))]
lmas<-merge(lmas,viol,by=c("PWS.ID","PWS.Name","PWS.Type","Primary.Source","Counties.Served","Cities.Served","PopulationServed.Count","Number.of.Facilities"),all.x=TRUE)

lmas<-lmas %>% 
   filter(is.na(Primary.Source)|Primary.Source=="Surface water",
          !is.na(LAKE_HISTORY_ID)) %>% 
  #,          Facility.Activity=="Active"
#          Availability.Description=="Permanent")
  select(LAKE_HISTORY_ID,waterbody,PWS.ID,PWS.Name,PWS.Type,Primary.Source,Counties.Served,Cities.Served,PopulationServed.Count,Number.of.Facilities,
         Owner.Type) %>% distinct() %>% 
  arrange(desc(PopulationServed.Count))


#write.csv(lmas,file="PWS.for.TMDL.and.sdwis.csv",row.names=FALSE)
DT::datatable(lmas, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```


## Int. PWS Lakes

Lakes with public water supply intakes that we are considering for intensive sampling. TO be clear, there are multiple intakes on many of these waterbodies.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
#merging lake id with the DOH ID

rm(list=ls())
#read in data
#Script to read in new database
library(tidyverse)
# library(ggmap)
# library(ggrepel)
# library(rgdal)
# library(sp)
# library(spatialEco)
# library(tidyverse)
library(knitr)
#Reading in new database
#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
setwd("C:/Users/amonion/New York State Office of Information Technology Services/BWAM - Lakes Database/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
lake<-read.csv("new_database/L_LAKE.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)

#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")
setwd("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization")



lake<-lake %>% select(LAKE_HISTORY_ID,PWS) %>% distinct() %>% 
  mutate(PWS=toupper(PWS)) %>% 
  filter(PWS=="YES")
locs<-location %>% select(LAKE_HISTORY_ID,LOCATION_HISTORY_ID,LOCATION_PWL_ID) %>% distinct()
pws<-merge(lake,locs,by=c('LAKE_HISTORY_ID'),all.x=TRUE)

#draft<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/2022_stayCALM_ponded_internal-draft-assessments.csv")
draft<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/2022_stayCALM_ponded_internal-draft-assessments.csv")

draft<-draft %>% 
  select(seg_id:parameter) %>% 
  distinct() %>% 
  rename(LOCATION_PWL_ID=seg_id) %>% 
  mutate(value=1)
draft<-merge(pws,draft,by=c('LOCATION_PWL_ID'),all.x=TRUE)
#restrict to public water supply use
draft<-draft %>% 
  filter(use=="source_of_water_supply") %>% 
  select(-LOCATION_HISTORY_ID) %>% distinct()



#restrict to those that have confirmed assessments
draft<-draft %>% 
  filter(grepl("unconfirmed",use_assessment)) %>% 
  distinct()

#Remove NYC reservoirs
draft<-draft %>% filter(!(LAKE_HISTORY_ID %in% c('1302WES0067','1302BOY0076','1302EAS0089','1302BOG0086','1302MID0062','1302DIV0083','1302CRO0059',
                                                 '1302TIT0103','1302CRO0109','1302AMA0050','1302MUS0044A','1302NEW0044',
                                                 '1702KEN1063','1302GIL0061','1302GLE0074','1302KIR0052','1404CAN0402A','1403PEP0358A',
                                                 '1306RON0815A','1402NEV0058B','1307ASH0848','1202SCH0638A','1301JER1042',
                                                 '1702HIL1052')))


#tally results
junk<-draft %>% 
  select(LAKE_HISTORY_ID,use_assessment) %>% 
  mutate(sampling="intensive") %>% 
  group_by(sampling) %>% 
  summarize(n=n()) %>% 
  ungroup()
library(kableExtra)
kable(junk)
```



```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
sites<-draft %>% select(LOCATION_PWL_ID,use_assessment)
#Reading in new database
#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/Current")
setwd("C:/Users/amonion/New York State Office of Information Technology Services/BWAM - Lakes Database/Current")
location<-read.csv("new_database/L_LOCATION.csv",na.strings=c("","NA"), stringsAsFactors=FALSE)
#setwd("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization")
setwd("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization")
coordinates<-location %>% filter(LOCATION_TYPE=="CENTROID"|LOCATION_NAME=="CHAUTAUQUA S CSLAP SITE 33.1") %>% select(LOCATION_PWL_ID,LOCATION_X_COORDINATE,LOCATION_Y_COORDINATE) %>% distinct()
sites<-merge(sites,coordinates,by=c('LOCATION_PWL_ID'),all.x = TRUE)

library(ggmap)
library(ggrepel)

nybox<-make_bbox(sites,lon=LOCATION_X_COORDINATE,lat=LOCATION_Y_COORDINATE)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites,aes(x=LOCATION_X_COORDINATE,y=LOCATION_Y_COORDINATE,label=LOCATION_PWL_ID),size=4))



```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
doh<-read.csv("data.requests/PWS_IDs.csv")
doh<-doh %>%
  select(LOCATION_PWL_ID,LAKE_HISTORY_ID,PWS.ID) %>% distinct() 
draft<-merge(draft,doh,by=c('LAKE_HISTORY_ID','LOCATION_PWL_ID'),all.x=TRUE)


```


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
lmas<-draft
#read in DWSP2 files
#wss<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/water.systems.summary.csv")
wss<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/water.systems.summary.csv")

lmas<-merge(wss,lmas,by=c('PWS.ID'),all=TRUE)

#too many individual facilities to be useful
# rep<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/facility.report.csv")
# lmas<-merge(rep,lmas,by=c('PWS.ID','PWS.Name','Primary.Source','PWS.Type','Owner.Type'),all=TRUE)

#viol<-read.csv("C:/Users/leneo/Dropbox/Alene/Rscripts/PWS_Prioritization/Water System Summary viol.csv")
viol<-read.csv("C:/Users/amonion/OneDrive - New York State Office of Information Technology Services/Rscripts/PWS_Prioritization/Water System Summary viol.csv")

#colnames(viol)[(colnames(viol) %in% colnames(lmas))]
lmas<-merge(lmas,viol,by=c("PWS.ID","PWS.Name","PWS.Type","Primary.Source","Counties.Served","Cities.Served","PopulationServed.Count","Number.of.Facilities"),all.x=TRUE)

lmas<-lmas %>% 
   filter(is.na(Primary.Source)|Primary.Source=="Surface water",
          !is.na(LAKE_HISTORY_ID)) %>% 
  #,          Facility.Activity=="Active"
#          Availability.Description=="Permanent")
  select(LAKE_HISTORY_ID,waterbody,PWS.ID,PWS.Name,PWS.Type,Primary.Source,Counties.Served,Cities.Served,PopulationServed.Count,Number.of.Facilities,
         Owner.Type) %>% distinct() %>% 
  arrange(desc(PopulationServed.Count))


#write.csv(lmas,file="PWS.for.TMDL.and.sdwis.csv",row.names=FALSE)
DT::datatable(lmas, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```


## Sc. PWS Streams

Flowing waters with public water supply intakes that we are intending to screen 2022-23. TO be clear, there are multiple intakes on many of these waterbodies.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
rm(list=ls())
sites<-read.csv("SMAS/Unassessed_PWS_Intakes_W_PWL_ID_From_Final_List_created_2022_02_15.csv")
sites<-sites %>% 
  rename(PWS.ID=WATER_SYSTEMID,
         PWS.Name=WATER_SYSTEM_NAME,
         Primary.Source=SOURCE_NAME,
         PWS.Type=FACILITY_TYPE,
         PopulationServed.Count=PWS_INTAKE_LAYER_POPULATION_SERVED) %>% 
  mutate(sampling="screening") %>% 
  distinct(sampling,PWS.ID,PWS.Name,PWS.Type,PopulationServed.Count,WATER_SYSTEM_TYPE,PWL_ID,WIN,NAME,DESCRIPT,PARTITEM,CLASSIFICA,Lat,Long) %>% distinct()
  

#tally results
junk<-sites %>% 
  select(PWL_ID,NAME,sampling) %>% distinct() %>% 
  group_by(sampling) %>% 
  summarize(n=n()) %>% 
  ungroup()
library(kableExtra)
kable(junk)

nybox<-make_bbox(sites,lon=Long,lat=Lat)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites,aes(x=Long,y=Lat,label=NAME),size=4))



DT::datatable(sites, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```

## Int. PWS Streams

Flowing waters with public water supply intakes that we are considering for intensive sampling. TO be clear, there are multiple intakes on many of these waterbodies.

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
rm(list=ls())
sites<-read.csv("SMAS/Unconfirmed_PWS_Intakes_W_PWL_ID_From_Final_List_created_2022_02_15.csv")
sites<-sites %>% 
  rename(PWS.ID=WATER_SYSTEMID,
         PWS.Name=WATER_SYSTEM_NAME,
         Primary.Source=SOURCE_NAME,
         PWS.Type=FACILITY_TYPE,
         PopulationServed.Count=PWS_INTAKE_LAYER_POPULATION_SERVED) %>% 
  mutate(sampling="intensive") %>% 
  distinct(sampling,PWS.ID,PWS.Name,PWS.Type,PopulationServed.Count,WATER_SYSTEM_TYPE,PWL_ID,WIN,NAME,DESCRIPT,Lat,Long) %>% distinct()
  

#tally results
junk<-sites %>% 
  select(PWL_ID,NAME,sampling) %>% distinct() %>% 
  group_by(sampling) %>% 
  summarize(n=n()) %>% 
  ungroup()
library(kableExtra)
kable(junk)

nybox<-make_bbox(sites,lon=Long,lat=Lat)

print(ny.map1<-qmap(nybox,source="osm",maptype="terrain",color="bw")+
  geom_point(data=sites,aes(x=Long,y=Lat,label=NAME),size=4))



DT::datatable(sites, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```