---
title: "PWS_sampling_strategy_2024"
author: "Giancarlo Coppola"
date: '`r paste("DRAFT Report Generated on:", format(Sys.Date(),format="%B %d, %Y"))`'
output: html_document
---

#  {.tabset}

```{r, include=FALSE, results='asis', message=FALSE, warning=FALSE}
options(scipen=999) # prevents scientific notation of site code

# filter out HAB reports --------------------------
pws_total<-newdata %>% 
  filter(PWS=="YES") %>% 
  mutate(year = substr(SAMPLE_DATE,1,4))

non_hab<-pws_total %>% filter(is.na(HS_HAB_STATUS),
                              SAMPLE_TYPE!='HAB')

# separate out sampled vs unsampled PWS
pws_sampled<-non_hab %>% filter(!is.na(SAMPLE_DATE))
pws_unsampled<-pws_total %>% filter(!LAKE_HISTORY_ID %in% pws_sampled$LAKE_HISTORY_ID)


# temp<-newdata %>% 
#   filter(PWS=="YES") %>% 
#   select(LAKE_HISTORY_ID,LOCATION_PWL_ID) %>% distinct() 
  
#newdata$year<-substr(newdata$SAMPLE_DATE,1,4)

# assessments havent been completed yet for 2023 so Im doing this manually
ponded_2023<-pws_sampled %>% 
  mutate(new_2023 = case_when(
    LAKE_HISTORY_ID %in% pws_sampled[pws_sampled$year !='2023',]$LAKE_HISTORY_ID ~ 'NO',
    TRUE ~ 'YES'
  )) %>% 
  filter(year=='2023', SAMPLE_TYPE == 'WATER COLUMN', DATA_PROVIDER == 'LCI') %>% 
  select(LOCATION_PWL_ID, LAKE_HISTORY_ID) %>% distinct()

# prepare assessment data
att_raw<-read.csv("C:/Users/gxcoppol/New York State Office of Information Technology Services/LMAS - PWS/PWS_Prioritization/2022.Data/2022_lmas_evaluations.csv")%>%
  rename(LOCATION_PWL_ID=seg_id) %>% distinct()

att<-att_raw %>% 
  filter(LOCATION_PWL_ID %in% pws_sampled$LOCATION_PWL_ID) %>% 
  left_join(pws_sampled %>% select(LOCATION_PWL_ID, LAKE_HISTORY_ID) %>% distinct(), by = 'LOCATION_PWL_ID')


# att_primer<-merge(temp,att_raw,by=c('LOCATION_PWL_ID'),all.x=TRUE) %>% 
#   filter(!(LAKE_HISTORY_ID %in%
#                         c('1302WES0067','1302BOY0076','1302EAS0089','1302BOG0086',
#                           '1302MID0062','1302DIV0083','1302CRO0059',
#                           '1302TIT0103','1302CRO0109','1302AMA0050','1302MUS0044A','1302NEW0044',
#                           '1702KEN1063','1302GIL0061','1302GLE0074','1302KIR0052','1404CAN0402A',
#                           '1403PEP0358A','1306RON0815A','1402NEV0058B','1307ASH0848','1202SCH0638A',
#                           '1301JER1042','1702HIL1052',
#                           # lakes removed for reasons described in the unassessed without data section
#                           '1501STA1011','1301TIO0152','1301SUM0193',
#                           '1311THO0274','0905CRA0309','1307KIN0837A','1307KIN0837','1307KIN0838A',
#                           '1301BRO0155',
#                           '0402RUS0033','1702BAR1108','1000CHA0001','0300ONT0000',
#                           '1005DOL0424','1301LUS0204','0302COB0118','1202TAN0655A','1202TAN0655B',
#                           '1303UPP0223',
#                           # lakes removed for reasons described in the unassessed with data section
#                           '0302SOD0096','0601OTS0404','0601OTS0404','0601OTS0404','0705SEN0369',
#                           '0705SEN0369','0705SEN0369','1201MAR0570','1301POP0191','1308TAG0869',
#                           '0705SEN0369','0705SEN0369','1301QUE0184A','1301STI0187A','1301STI0187A')))


# att<-att_primer %>% 
#   filter(!LAKE_HISTORY_ID %in% ponded_2023[ponded_2023$new_2023!='YES',]) %>%
#   group_by(LAKE_HISTORY_ID) %>% 
#   summarise(n_samples_collected = max(n_samples_collected)) %>% 
#   bind_rows(ponded_2023[ponded_2023$new_2023=='YES',])



# att_never_sampled<-att %>% filter(is.na(n_samples_collected))

# no PWL waterbodies removed for various reasons described in 2023 markdown
no_pwl_remove<-c('0202CHA0122','0705CAY0296','0302UWB0008','0601UWB0395','1003RAY0091',
                 '0501ELM0019','1301UWB0413', '1306UWB0683','1101DOR0098',
                 '1303UPP0223','0503HOR0035','1101UWB1081','1201UWB0662','1306GLE0668','1306MIL0515',
                 '0100ERI5000', # lake Erie - have to remove these and put justification in markdown
                 '1202TAN0655D','1202UWB0589A','1501UWB1007A', # have data - have to remove these and put justification in markdown
                 '1302WES0067','1302BOY0076','1302EAS0089','1302BOG0086', # everything below has been copy pasted from above
                   '1302MID0062','1302DIV0083','1302CRO0059',
                   '1302TIT0103','1302CRO0109','1302AMA0050','1302MUS0044A','1302NEW0044',
                   '1702KEN1063','1302GIL0061','1302GLE0074','1302KIR0052','1404CAN0402A',
                   '1403PEP0358A','1306RON0815A','1402NEV0058B','1307ASH0848','1202SCH0638A',
                   '1301JER1042','1702HIL1052',
                   # lakes removed for reasons described in the unassessed without data section
                   '1501STA1011','1301TIO0152','1301SUM0193',
                   '1311THO0274','0905CRA0309','1307KIN0837A','1307KIN0837','1307KIN0838A', '1301BRO0155',
                   '0402RUS0033','1702BAR1108','1000CHA0001','0300ONT0000',
                   '1005DOL0424','1301LUS0204','0302COB0118','1202TAN0655A','1202TAN0655B',
                   '1303UPP0223',
                   # lakes removed for reasons described in the unassessed with data section
                   '0302SOD0096','0601OTS0404','0601OTS0404','0601OTS0404','0705SEN0369','0705SEN0369','0705SEN0369','1201MAR0570','1301POP0191','1308TAG0869',
                   '0705SEN0369','0705SEN0369','1301QUE0184A','1301STI0187A','1301STI0187A')


doh<-read.csv("C:/Users/gxcoppol/New York State Office of Information Technology Services/LMAS - PWS/PWS_Prioritization/data.requests/PWS_IDs.csv") %>%
  select(LOCATION_PWL_ID,LAKE_HISTORY_ID,PWS.ID) %>% distinct() 

# filter out lakes that were removed because of specific reasons 
lo<-location %>% filter(LAKE_HISTORY_ID %in% bwrm_check$LAKE_HISTORY_ID,
                        LOCATION_HISTORY_ID !='1301UWB0438P_BEACH')

bwrm_check<-att_never_sampled %>% filter(!LAKE_HISTORY_ID %in% no_pwl_remove) %>% 
  left_join(lake, by = 'LAKE_HISTORY_ID') %>% 
  left_join(doh, by = 'LAKE_HISTORY_ID') #%>% 
  #left_join(lo, by = 'LAKE_HISTORY_ID')

write.csv(bwrm_check, 'C:/Users/gxcoppol/New York State Office of Information Technology Services/LMAS - PWS/PWS_Prioritization/bwrm_check.csv', row.names = F, na = '')

nopwl<-att %>% filter(is.na(LOCATION_PWL_ID))
att<-att %>% filter(!is.na(LOCATION_PWL_ID))



```


## Assessed PWS

PWS that have sufficient data for an assessment (n_years_sampled >=2; n_samples>=6; use== source_of_water_supply)

Because there are multiple SDWIS ids per lake (multiple intakes per lake), we reported this list in two tables:

1. PWS lakes that can be assessed

2. Data from table 1 + SDWIS id 


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

table1_expanded<-att %>% 
  filter(n_years_sampled >=2| n_samples_collected>=6) %>% 
  filter(use== 'source_of_water_supply',
         within_period=='TRUE',
         !LAKE_HISTORY_ID %in% no_pwl_remove) %>% 
  select(LAKE_HISTORY_ID, LOCATION_PWL_ID) %>% distinct()

table1b_PWS_info<-merge(table1_expanded, doh, by = c('LAKE_HISTORY_ID','LOCATION_PWL_ID'), all.x = T)

DT::datatable(table1_expanded, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(table1b_PWS_info, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```

## Unassessed PWS with Data

PWS that have data but not enough for an assessment.

Because there are multiple SDWIS ids per lake (multiple intakes per lake), we reported this list in two tables:

1. PWS lakes with data

2. Data from table 1 + SDWIS id 


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

table2<-pws_sampled %>% 
  filter(!LAKE_HISTORY_ID %in% c(table1_expanded$LAKE_HISTORY_ID,
                                no_pwl_remove)) %>% 
  select(LAKE_HISTORY_ID, LOCATION_PWL_ID) %>% distinct()

table2b<-table2 %>% left_join(doh, by = c('LAKE_HISTORY_ID','LOCATION_PWL_ID'))

DT::datatable(table2, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(table2b, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```



## Unassessed PWS without Data

PWS that were never sampled.

Because there are multiple SDWIS ids per lake (multiple intakes per lake), we reported this list in two tables:

1. PWS lakes with data

2. Data from table 1 + SDWIS id 


```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

table3<-pws_unsampled %>% 
  filter(!LAKE_HISTORY_ID %in% c(no_pwl_remove)) %>% 
  bind_rows(pws_sampled[pws_sampled$LAKE_HISTORY_ID %in% c('0705UWB0340A','1301FER0028C'),]) %>% 
  select(LAKE_HISTORY_ID, LOCATION_PWL_ID) %>% 
  distinct()

table3b<-table3 %>% left_join(doh, by = c('LAKE_HISTORY_ID'))
fucked<-table3b %>% filter(!LAKE_HISTORY_ID %in% bwrm$LAKE_HISTORY_ID)
fucked2<-bwrm %>% filter(!LAKE_HISTORY_ID %in% table3$LAKE_HISTORY_ID)

fdat<-newdata %>% filter(LAKE_HISTORY_ID %in% fucked$LAKE_HISTORY_ID)


DT::datatable(table1_expanded, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
DT::datatable(table1b_PWS_info, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))

```
